Constraining gravitational wave amplitude birefringence with GWTC-3

### 2023-03-01 src/tex/ms.tex
section rearrangement
Add content of GW200129

The two events that show the largest tension with $\kappa = 0$ is GW170818, for which $|\mu_i / \sigma_i| = \variable{output/GW170818_constraint.txt}$, and GW200129\_065458 (henceforth GW200129), for which $|\mu_i / \sigma_i| = \dots$.
\mi{THOMAS: please add the number for GW200129. Also, we don't need that many significant figures, one decimal will do.}
However, as we discuss further below, we have reason to think that the GW200129 preference for $\kappa < 0$ might be driven by noise anomalies in the Virgo detector; for that reason we consider the effect of excluding it from the joint result below.

@@

\begin{figure}
    \script{posterior_kappa.py}
    \includegraphics[width=\columnwidth]{figures/posterior_kappa.pdf}
    \caption{
        The generic and restricted posterior of $\kappa$.
        The blue solid line shows the restricted posterior of $\kappa$, and the blue dashed lines enclose the $90\%$ confidence interval.
        The orange solid line shows the generic posterior of $\kappa$, and the orange dashed lines enclose the $90\%$ confidence interval.
        The dashed line marks the absence of birefringence ($\kappa=0$).
        \mi{THOMAS: please add a second set of curves \emph{without} GW200129; use the same colors you have now but plot them with a dashed trace}
    }
    \label{fig:posterior_kappa}
\end{figure}

% Case study

% Case: GW170818
\subsection{Notable events}

@@

\subsubsection{GW170818}
GW170818 gave the largest tension with $\kappa=0$, when judged by $\mu_i/\sigma_i$ in Fig.~\ref{fig:violin_kappa}.
Figure \ref{fig:corner_GW170818} shows the birefringence effect allows a region of $\kappa>0$ for
 larger distance and a smaller inclination angles than would be allowed in the absence of birefringence.
We can make sense of this result because a smaller inclination angle will make the amplitude ratio between the two circular polarizations smaller.
Then the birefringence effect will increase the amplitude ratio during the propagation.
The larger distance is for compensating the combined effect of a nonzero $\kappa$ and a smaller inclination angle.
We can see this effect in other events as well, but not as strongly as in GW170818.

@@

\subsubsection{GW200129}
\label{sec:GW200129}

@@

@@

% Case: GW190521 (massive BBH)
\subsubsection{Result on GW190521}
Even with the frequency-dependent birefringence model, some events still have a bimodal $\kappa$ distribution.
Consider GW190521, the most massive binary black hole merger in the events we included.
The degeneracy between $\kappa$ and $\iota$ cannot be broken by the frequency-dependent birefringence model, as shown in Fig.~\ref{fig:corner_GW190521}.
The reason could be this GW event's frequency range is too narrow to break the degeneracy.
The effect of birefringence at different frequencies within the range is similar, so the degeneracy cannot be broken.
Therefore, the narrow frequency range can be a possible reason why the $\kappa$ distribution of GW190521 is bimodal.
Note that $\kappa=0$ is still within the $90\%$ confidence interval, with a non-negligible probability.
The bimodal $\kappa$ distribution only means that the degeneracy between $\kappa$ and $\iota$ cannot be broken in this case.

% Case: GW200129 (glitch in Virgo data)
\section{Discussion}
\label{sec:Discussion}

>>>>

The two events that show the largest tension with $\kappa = 0$ is GW170818, for which $|\mu_i / \sigma_i| =$ \variable{output/GW170818_constraint.txt}, and GW200129\_065458 (henceforth GW200129), for which $|\mu_i / \sigma_i| =$ \variable{output/GW200129_constraint.txt}.
However, as we discuss further below, we have reason to think that the GW200129 preference for $\kappa < 0$ might be driven by noise anomalies in the Virgo detector; for that reason we consider the effect of excluding it from the joint result below.

@@

\begin{figure}
    \script{posterior_kappa.py}
    \includegraphics[width=\columnwidth]{figures/posterior_kappa.pdf}
    \caption{
        The generic and restricted posterior of $\kappa$.
        The blue solid line shows the restricted posterior of $\kappa$, and the blue dashed lines enclose the $90\%$ confidence interval.
        The orange solid line shows the generic posterior of $\kappa$, and the orange dashed lines enclose the $90\%$ confidence interval.
        The dashed line marks the absence of birefringence ($\kappa=0$).
    }
    \label{fig:posterior_kappa}
\end{figure}

% Case studies
\subsection{Notable events}

@@

% Case: GW170818
\subsubsection{GW170818}
GW170818 gave the largest tension with $\kappa=0$, when judged by $\mu_i/\sigma_i$ in Fig.~\ref{fig:violin_kappa}.
Figure \ref{fig:corner_GW170818} shows the birefringence effect allows a region of $\kappa>0$ for
 larger distance and a smaller inclination angles than would be allowed in the absence of birefringence.
We can make sense of this result because a smaller inclination angle will make the amplitude ratio between the two circular polarizations smaller.
Then the birefringence effect will increase the amplitude ratio during the propagation.
The larger distance is for compensating the combined effect of a nonzero $\kappa$ and a smaller inclination angle.
We can see this effect in other events as well, but not as strongly as in GW170818.

@@

% Case: GW200129 (glitch in Virgo data)
\subsubsection{GW200129}
\label{sec:GW200129}

@@

% Case: GW190521 (massive BBH)
\subsubsection{GW190521}
Even with the frequency-dependent birefringence model, some events still have a bimodal $\kappa$ distribution.
Consider GW190521, the most massive binary black hole merger in the events we included.
The degeneracy between $\kappa$ and $\iota$ cannot be broken by the frequency-dependent birefringence model, as shown in Fig.~\ref{fig:corner_GW190521}.
The reason could be this GW event's frequency range is too narrow to break the degeneracy.
The effect of birefringence at different frequencies within the range is similar, so the degeneracy cannot be broken.
Therefore, the narrow frequency range can be a possible reason why the $\kappa$ distribution of GW190521 is bimodal.
Note that $\kappa=0$ is still within the $90\%$ confidence interval, with a non-negligible probability.
The bimodal $\kappa$ distribution only means that the degeneracy between $\kappa$ and $\iota$ cannot be broken in this case.

@@

\section{Discussion}
\label{sec:Discussion}
